---
title: "615 Assignment Strawberry 3"
author: "Yiming Chen"
date: "2024-10-21"
output: pdf_document
---
#Preparing data for analysis —— Strawberries

##read and explore the data
```{r}
library(knitr)
library(kableExtra)
library(tidyverse)
library(dplyr)
library(readr)
library(tidyr)
library(stringr)
library(ggplot2)
```

Read in the dataset and take a first look.
```{r}
strawberry <- read_csv("strawberries25_v3.csv", col_names = TRUE)
head(strawberry)
```

Remove the (D) term in Value and CV% columns
```{r}
strawberry <- strawberry %>%
  mutate(
    Value = ifelse(Value == "(D)", NA, Value),
    `CV (%)` = ifelse(`CV (%)` == "(D)", NA, `CV (%)`)
  )
```

Do data cleaning for the Domain column, rearrange the info in this column into three columns: chemical category, name and number
```{r}
strawberry <- strawberry %>%
  mutate(Category = case_when(
    Domain == "Total" ~ NA_character_,
    str_detect(Domain, "CHEMICAL") ~ str_trim(str_remove(Domain, "CHEMICAL, ")),
    TRUE ~ Domain
  ))
unique(strawberry$Category)
```
Clean and transform the strawberry dataset by creating new columns based on specific conditions and regex extraction from the Domain Category field while handling special cases, such as missing values or unspecified categories.
```{r}
strawberry <- strawberry %>%
  mutate(
    Name = case_when(
      Category == "TOTAL" ~ NA_character_,
      str_detect(`Domain Category`, fixed(Category)) & str_detect(`Domain Category`, "\\(.*=.*\\)") ~ 
        str_extract(`Domain Category`, "(?<=\\().*?(?=\\s?=)"),
      str_detect(`Domain Category`, fixed(Category)) & str_detect(`Domain Category`, "\\(.*\\)") ~ 
        str_extract(`Domain Category`, "(?<=\\().*?(?=\\))"),
      TRUE ~ NA_character_
    ),
    Number = case_when(
      Category == "TOTAL" ~ NA_real_,
      str_detect(`Domain Category`, fixed(Category)) & str_detect(`Domain Category`, "\\(.*=.*\\)") ~ 
        as.numeric(str_extract(`Domain Category`, "(?<=\\=\\s?).*?(?=\\))")),
      str_detect(`Domain Category`, fixed(Category)) & str_detect(`Domain Category`, "\\(.*\\)") ~ 
        NA_real_,
      TRUE ~ NA_real_
    )
  )

strawberry <- strawberry %>%
  mutate(Category = case_when(
    `Domain Category` == "NOT SPECIFIED" ~ NA_character_,
    TRUE ~ Category
  ))
```
data cleaning for AREA GROWN, the numerical intervals of the planted area are reintegrated inside the new columns, respectively, with the column names of Min and Max
```{r}
strawberry <- strawberry %>%
  mutate(
    Min = case_when(
      str_detect(Name, "100 OR MORE ACRES") ~ 100,
      str_detect(Name, "TO") ~ as.numeric(str_extract(Name, "^[0-9.]+")),
      TRUE ~ NA_real_
    ),
    Max = case_when(
      str_detect(Name, "100 OR MORE ACRES") ~ "MORE",
      str_detect(Name, "TO") ~ str_extract(Name, "(?<=TO )^[0-9.]+"),
      TRUE ~ NA_character_
    )
)
```

Create a new column 'Unit' by extracting the substring after 'MEASURED'.
Create a new column 'Type' by extracting either 'BEARING' or 'ORGANIC'.
Create a new column 'Operation' by extracting the remaining parts of the string,
Removing the 'MEASURED' part, the Unit and the Type, keeping the rest.
Create a new column 'Operation' by extracting the remaining parts of the string,
Removing the 'MEASURED', 'BEARING', 'ORGANIC', and 'STRAWBERRIES' parts.

```{r}
strawberry <- strawberry %>%
  mutate(Unit = str_extract(strawberry$`Data Item`, "(?<=MEASURED ).*"))

strawberry <- strawberry %>%
  mutate(Type = str_extract(strawberry$`Data Item`, "BEARING|ORGANIC"))

strawberry <- strawberry %>%
  mutate(Operation = str_replace_all(strawberry$`Data Item`, "MEASURED.*|BEARING|ORGANIC", "") %>%
           str_trim())

strawberry <- strawberry %>%
  mutate(Operation = str_replace_all(strawberry$`Data Item`, "MEASURED.*|BEARING|ORGANIC|STRAWBERRIES(, | - )", "") %>%
           str_replace_all("[-,]", "") %>%
           str_trim())
```

Export the cleaned dataset as a CSV file.
```{r}
write.csv(strawberry, "cleaned_strawberries.csv", row.names = FALSE)
```

##EDA
Check data types
```{r}
str(strawberry)
```

Convert 'Value' to numeric, removing non-numeric characters.
Convert 'CV (%)' to numeric, removing non-numeric characters (including %, parentheses).
Check if conversion was successful.
Check for any NAs introduced after conversion.
Summary statistics for 'Value' and 'CV (%)'.
Check for missing values in 'Value' and 'CV (%)'.
Histogram for 'CV (%)'.
```{r}
strawberry$Value <- as.numeric(gsub("[^0-9.]", "", strawberry$Value))

strawberry$`CV (%)` <- as.numeric(gsub("[^0-9.]", "", strawberry$`CV (%)`))

str(strawberry$Value)
str(strawberry$`CV (%)`)

sum(is.na(strawberry$Value))
sum(is.na(strawberry$`CV (%)`))

summary(strawberry$Value)
summary(strawberry$`CV (%)`)

sum(is.na(strawberry$Value))
sum(is.na(strawberry$`CV (%)`))

ggplot(strawberry, aes(x = `CV (%)`)) +
  geom_histogram(binwidth = 1, col = "pink", fill = "pink") +
  labs(title = "Distribution of CV (%)", x = "CV (%)", y = "Frequency")
```

The Value column shows a strong right skew with most data concentrated at lower values and only a few larger ones.
The CV (%) column displays a more spread distribution. The frequent occurrence of CV values between 20% and 30% may indicate that this range represents the typical variation in the dataset. However, the existence of high CV values suggests that certain categories or items show much higher variability.

Bar plot for 'Type' column
```{r}
ggplot(strawberry, aes(x=Type)) +
  geom_bar(fill="lightgreen") +
  theme(axis.text.x = element_text(angle=45, hjust=1)) +
  labs(title="Distribution of Type")
```
The BEARING type is the most common category in the Type column, while ORGANIC data points are minimal.
The significant proportion of NA values suggests that a substantial amount of Type information is missing, which could have implications for further analyses or interpretations related to strawberry types.


```{r}
library(tidyverse)
library(ggplot2)
strawberry = read.csv("cleaned_strawberries.csv")
view(strawberry)
```

Count the total number of occurrences of chemicals in each category.
Create a bar chart for the total counts of each category.
```{r}
filtered_data <- strawberry %>%
  filter(State == "FLORIDA" & 
         Category %in% c("FUNGICIDE", "OTHER", "HERBICIDE", "INSECTICIDE"))

category_total_counts <- filtered_data %>%
  group_by(Category) %>%
  summarise(Total_Count = n()) %>%
  arrange(desc(Total_Count))

ggplot(category_total_counts, aes(x = reorder(Category, -Total_Count), y = Total_Count, fill = Category)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = c("FUNGICIDE" = "skyblue", "HERBICIDE" = "lightgreen",
                               "INSECTICIDE" = "orange", "OTHER" = "purple")) +  # Custom colors for each category
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5, face = "bold"),  # Style x-axis labels
        axis.text = element_text(size = 12, color = "darkgray"),  # Adjust font size and color
        plot.title = element_text(hjust = 0.5, face = "bold", color = "darkblue"),  # Center and style title
        panel.grid.major = element_line(color = "lightgray", linetype = "dashed")) +  # Dashed grid lines
  labs(title = "Total Count of Chemicals by Category in Florida",
       x = "Category", y = "Total Count")
```
This chart effectively highlights the distribution of chemical categories in Florida, making it easy to compare the total counts visually.The bar chart displays the counts for each category in Florida, showing "FUNGICIDE" and "INSECTICIDE" as the most frequently occurring categories, with "OTHER" and "HERBICIDE" following.


Filter data to include only the categories FUNGICIDE, OTHER, HERBICIDE, INSECTICIDE, 
State = New York, and Program = SURVEY.
Count the number of occurrences of each chemical name within each category.
Create a function to plot bar chart for each category.
Check if there's data to plot.
Generate and print plots for each category.
Print only plots that were successfully created.
```{r,warning=FALSE,message=FALSE}
filtered_data <- strawberry %>%
  filter(State == "FLORIDA" & Program == "SURVEY" & 
         Category %in% c("FUNGICIDE", "OTHER", "HERBICIDE", "INSECTICIDE"))

category_chemical_counts <- filtered_data %>%
  group_by(Category, Name) %>%
  summarise(Count = n()) %>%
  arrange(Category, desc(Count))

plot_category <- function(category_name) {
  subset_data <- category_chemical_counts %>%
    filter(Category == category_name)
  
  if(nrow(subset_data) == 0) {
    message(paste("No data available for category:", category_name))
    return(NULL)
  }
  
  ggplot(subset_data, aes(y = reorder(Name, Count), x = Count, fill = Count)) +
    geom_bar(stat = "identity", width = 0.7) +
    scale_fill_gradient(low = "skyblue", high = "darkblue") +  # Add a gradient fill
    theme_minimal(base_size = 14) +
    theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5, face = "italic"),  # Style y-axis labels
          axis.text = element_text(size = 10, color = "darkgray"),  # Adjust font size and color
          plot.title = element_text(hjust = 0.5, face = "bold", color = "darkblue"),  # Center and style title
          panel.grid.major = element_line(color = "lightgray", linetype = "dashed")) +  # Dashed grid lines
    labs(title = paste("Counts of Chemicals for", category_name, "in Florida"),
         y = "Chemical Name", x = "Count")
}

categories <- c("FUNGICIDE", "HERBICIDE", "INSECTICIDE", "OTHER")
plots <- lapply(categories, plot_category)

for (plot in plots) {
  if (!is.null(plot)) {
    print(plot)
  }
}

```
The resulting plots display each category’s chemical counts in Florida with a color gradient indicating the count levels, making it easy to compare the usage frequency of each chemical within each category.
These visualizations are useful for assessing which chemicals are most common within each category in Florida’s survey data, especially with the gradient highlighting relative frequencies. 


```{r}
filtered_data <- strawberry %>%
  filter(State == "CALIFORNIA" & 
         Category %in% c("FUNGICIDE", "OTHER", "HERBICIDE", "INSECTICIDE"))

# Count the total number of occurrences of chemicals in each category
category_total_counts <- filtered_data %>%
  group_by(Category) %>%
  summarise(Total_Count = n()) %>%
  arrange(desc(Total_Count))

# Create a bar chart for the total counts of each category
ggplot(category_total_counts, aes(x = reorder(Category, -Total_Count), y = Total_Count, fill = Category)) +
  geom_bar(stat = "identity", width = 0.7) +
  scale_fill_manual(values = c("FUNGICIDE" = "skyblue", "HERBICIDE" = "lightgreen",
                               "INSECTICIDE" = "orange", "OTHER" = "purple")) +  # Custom colors for each category
  theme_minimal(base_size = 14) +
  theme(axis.text.x = element_text(angle = 0, hjust = 0.5, vjust = 0.5, face = "bold"),  # Style x-axis labels
        axis.text = element_text(size = 12, color = "darkgray"),  # Adjust font size and color
        plot.title = element_text(hjust = 0.5, face = "bold", color = "darkblue"),  # Center and style title
        panel.grid.major = element_line(color = "lightgray", linetype = "dashed")) +  # Dashed grid lines
  labs(title = "Total Count of Chemicals by Category in California",
       x = "Category", y = "Total Count")
```
This chart effectively highlights the distribution of chemical categories in California, making it easy to compare the total counts visually.The bar chart displays the counts for each category in California, showing "INSECTICIDE" and "FUNGICIDE" as the most frequently occurring categories, with "OTHER" and "HERBICIDE" following.

```{r}

# Filter data to include only the categories FUNGICIDE, OTHER, HERBICIDE, INSECTICIDE, 
# State = New York, and Program = SURVEY
filtered_data <- strawberry %>%
  filter(State == "CALIFORNIA" & Program == "SURVEY" & 
         Category %in% c("FUNGICIDE", "OTHER", "HERBICIDE", "INSECTICIDE"))

# Count the number of occurrences of each chemical name within each category
category_chemical_counts <- filtered_data %>%
  group_by(Category, Name) %>%
  summarise(Count = n()) %>%
  arrange(Category, desc(Count))

# Create a function to plot bar chart for each category
plot_category <- function(category_name) {
  subset_data <- category_chemical_counts %>%
    filter(Category == category_name)
  
  # Check if there's data to plot
  if(nrow(subset_data) == 0) {
    message(paste("No data available for category:", category_name))
    return(NULL)
  }
  
  ggplot(subset_data, aes(y = reorder(Name, Count), x = Count, fill = Count)) +
    geom_bar(stat = "identity", width = 0.7) +
    scale_fill_gradient(low = "skyblue", high = "darkblue") +  # Add a gradient fill
    theme_minimal(base_size = 14) +
    theme(axis.text.y = element_text(angle = 0, hjust = 1, vjust = 0.5, face = "italic"),  # Style y-axis labels
          axis.text = element_text(size = 10, color = "darkgray"),  # Adjust font size and color
          plot.title = element_text(hjust = 0.5, face = "bold", color = "darkblue"),  # Center and style title
          panel.grid.major = element_line(color = "lightgray", linetype = "dashed")) +  # Dashed grid lines
    labs(title = paste("Counts of Chemicals for", category_name, "in California"),
         y = "Chemical Name", x = "Count")
}

# Generate and print plots for each category
categories <- c("FUNGICIDE", "HERBICIDE", "INSECTICIDE", "OTHER")
plots <- lapply(categories, plot_category)

# Print only plots that were successfully created
for (plot in plots) {
  if (!is.null(plot)) {
    print(plot)
  }
}


```
The resulting plots display each category’s chemical counts in California with a color gradient indicating the count levels, making it easy to compare the usage frequency of each chemical within each category.
These visualizations are useful for assessing which chemicals are most common within each category in California’s survey data, especially with the gradient highlighting relative frequencies. 

A process to retrieve and display Global Harmonized System (GHS) hazard information for various chemicals.
GHS Search and Hazard Retrieval:
The code uses functions GHS_searcher and hazards_retriever to look up GHS hazard information based on a chemical identifier (result). This hazard information is saved in hazards.
Storing Results in a List:

The retrieved hazard data (hazards) for each chemical is stored in a list named results_list, with each chemical's name as the list key.
Converting to a Data Frame:

The results_list is converted to a data frame (results_df) using enframe, with the column names set to "Chemical" and "Hazard_Statements."
unnest is used to expand Hazard_Statements (which may contain multiple statements for each chemical) into separate rows.
Displaying the Data Frame:

The results_df is displayed, showing each chemical and its associated GHS hazard statements. Each hazard code and statement (e.g., "H302: Harmful if swallowed") is presented per row for easy reference.
```{r,warning=FALSE, message=FALSE}
library(tidyverse)
library(PubChemR)

GHS_searcher <- function(result_json_object) {
  hierarchies <- result_json_object[["result"]][["Hierarchies"]][["Hierarchy"]]
  
  for (i in seq_along(hierarchies)) {
    if (hierarchies[[i]][["SourceName"]] == "GHS Classification (UNECE)") {
      return(i)
    }
  }
  # Return NULL if GHS Classification is not found
  return(NULL)
}

hazards_retriever <- function(index, result_json_object) {
  if (is.null(index)) {
    return(NA)  # Return NA if GHS data is not available
  }
  
  hierarchy <- result_json_object[["result"]][["Hierarchies"]][["Hierarchy"]][[index]]
  nodes <- hierarchy[["Node"]]
  hazard_statements <- c()
  i <- 1
  
  while (i <= length(nodes) && str_detect(nodes[[i]][["Information"]][["Name"]], "^H")) {
    hazard_statements <- c(hazard_statements, nodes[[i]][["Information"]][["Name"]])
    i <- i + 1
  }
  if (length(hazard_statements) == 0) {
    return(NA)
  }
  return(hazard_statements)
}

# List of chemicals to process
chemical_vec <- c("reynoutria sachaline", "flutriafol", "chloropicrin")

# Initialize an empty list to store results
results_list <- list()

for (chemical in chemical_vec) {
  result <- get_pug_rest(
    identifier = chemical,
    namespace = "name",
    domain = "compound",
    operation = "classification",
    output = "JSON"
  )
  
  ghs_index <- GHS_searcher(result)
  hazards <- hazards_retriever(ghs_index, result)
  
  # Store the results in a list
  results_list[[chemical]] <- hazards
}

# Convert the results list into a data frame
results_df <- results_list %>%
  enframe(name = "Chemical", value = "Hazard_Statements") %>%
  unnest(cols = c(Hazard_Statements))

# Display the data frame
print(results_df)

```
The output table lists chemicals and their hazard statements. For example:
"flutriafol" has multiple hazards like "H302: Harmful if swallowed" and "H410: Toxic to aquatic life."
"chloropicrin" includes hazards such as "H330: Fatal if inhaled" and "H410: Very toxic to aquatic life."
This table provides a comprehensive overview of hazard classifications for each chemical, facilitating risk assessment and safety measures.







